---
title: "Nigeria Data Health Survey Analysis"
subtitle: "A look into the effect on Govt Policy in Nigeria Population Growth"
author: "Ian Zablon Owuor"
date: "2024-10-31"
output:
  html_document:
    theme: cosmo          # Clean, modern theme (options: cerulean, journal, flatly, etc.)
    highlight: tango      # Syntax highlighting style
    toc: true             # Include table of contents
    toc_float: true       # Floating TOC for easy navigation
    toc_depth: 3          # Depth of headers in TOC
    code_folding: show    # Allow users to toggle code visibility
    number_sections: true # Auto-number section headings
    css: custom.css       # Optional: link to custom CSS file for further styling
    df_print: paged       # Nicer table output for data frames
editor_options:
  markdown:
    wrap: 72
---

------------------------------------------------------------------------

------------------------------------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Over the past few years, Nigeria has recorded one of the highest
increase in population not only in Africa but in the whole world. An
increase in population leads to an increase in the workforce of a
country there-by causing an increase in the Gross Domestic Production.
In as much as a positive increase in the population causes an increase
in GDP, Population increase also comes with its own share of challenges.
Nigeria as a country has recorded an increased population over the years
and as at 2023, the population of Nigeria is 233.8 million people. By
2050, the Nigerian population is expected to double and by 2100,
Nigeria’s population will jump to 791 million, overtaking Chinaa to
become the second-most populous country in the world after only India.
This is something that can’t be neglected as more than 80 million
Nigerians already live below the poverty line ( Okpala, A. O.9(4),
63-77. ) The Federal Government of Nigeria has taken steps to ensure
that the population growth rate of Nigeria is dealt with. The minister
of health Osagie Ehanire said that one of the strategies that the
Federal Government has employed to curb the fast rate of population
growth is the development of a Task-Shifting/Task-sharing policy which
seeks to achieve rational redistribution of clinical and other tasks
among healthcare workers that will ensure an increase access to family
planning and sexual and reproductive health services especially to rural
and hard-reach areas. Another policy that was adopted by the Federal
Government of Nigeria is the reduction in the fertility from the levels
back then which were at 6 children/family to an average of 4
children/family. It also suggested that the optimum marriage age should
be 18 years for women and 24 years for men and advocated that the
pregnancies be limited to women between the age of 18-35 year range at
an interval of 2 years (Mazzocco, K.9(3), 1-12.) Finally the Federal
Government of Nigeria decided to increase the awareness of its people
through mass media forms particularly through radio to educate the
population about the need for family planning and maternal child health
programs. In this analysis we look at if and how these policies have
been fulfilled in the step to ensuring that the birth rate of Nigeria is
put under control.

## **Question 1.1**

### **Research question 1.1**

The Nigerian Federal Government has put policies to reduce the fast
increase in the population growth. Have these policies adopted by the
Federal Government of Nigeria helped to reduce the population growth
rate in the Country?. Some of the reasons that might have caused an
increase in population include. • Drop in the age of marriage for women
has led to increase in birth rate. • Reduction in the pregnancy age (\<
18) and older women (\> 35) getting pregnant. • Poor public education by
mass media particularly radio. • High illiteracy levels. • Religious
beliefs that encourage early birth in women. • Low levels of higher
education.

### **Research question 1.2**

**How does your research Question pertain to public policy.**

**1. Population Control Policies**: The research question directly
addresses the effectiveness of pop ulation control policies implemented
by the Federal Government of Nigeria. These policies are of significant
public interest and are aimed at managing the country’s demographic
growth. **2. Healthcare and Reproductive Services**: The policies are
related to healthcare services, especially family planning and maternal
health. They influence the availability and accessibility of such
services to the public, particularly women and families. **3. Education
and Awareness:** The research question touches on the impact of public
education efforts through mass media, such as radio. It evaluates the
role of education and awareness campaigns in shaping reproductive
behavior.

• Who are the primary Stakeholders or groups affected by this issue?

**1. The Nigerian Government**: The government plays a central role in
designing, implementing, and monitoring population control policies. It
is invested in the outcomes of these policies to achieve demographic and
economic goals. **2. Healthcare Providers:** Health workers and
organizations are essential stakeholders as they deliver family planning
and reproductive health services. Their workload may be affected by
changes in reproductive behavior. **3. Women and Families:** Women of
childbearing age, couples, and families are directly affected by the
policies that may influence their reproductive decisions. **4. Religious
and Cultural Institutions:** These institutions can play a role in
shaping cultural norms and beliefs about family planning and childbirth.
**5. Mass Media Organizations:** Radio and other mass media outlets are
involved in disseminating information about family planning and
reproductive health. They have a role in educating the public.

• In what ways can understanding or addressing this issue influence
policy decisions or outcomes?

**1. Policy Evaluation:** The research can provide evidence-based
insights into the effectiveness of existing policies. If the policies
are found to be successful, this can reinforce their continuation. If
they are ineffective, it may lead to policy adjustments or the
development of new strategies. **2. Resource Allocation:** The findings
can guide the allocation of resources to healthcare, education, and
awareness campaigns. Policymakers can prioritize investments in areas
that have a more significant impact on reducing birth rates. **3. Public
Awareness:** Understanding the role of mass media in shaping
reproductive behavior can lead to more targeted and effective public
awareness campaigns. This can improve access to information about family
planning and maternal health. **4. Cultural and Religious Sensitivity:**
Policymakers may need to consider cultural and religious beliefs when
designing and implementing policies. This research can shed light on how
these factors influence reproductive choices. **5. Gender Equality and
Education:** If a drop in the age of marriage for women or high
illiteracy levels is identified as a driver of population growth,
policies promoting gender equality and education may be more strongly
emphasized.

### **Question 1.3**

**Personal Reflection.**

• Share a bit of your perspective or experiences that led you to this
research question. My major concern for this research was the
sustainable development for Nigeria. The increase in population in
Nigeria has had significant environmental, social and economic impact
making access to basic health services, jobs and employment hard to come
by.

•Why does this questionn resonates with you personally and what drove
your interest to explore it further. My personal motivation to explore
this question further was driven by the desire to contribute to more
informed policy decisions, advocate for change and better understand the
complex factors that play in population growth in Nigeria.

## **Graded Task 2: Exploration**

The Demographic and Health Surveys (DHS) is a series of standardized
surveys conducted periodically in developing countries that includes
collecting data on various aspects of population, health, and
nutrition.These surveys are a gold-standard source of reliable public
health data for many countries and international bodies.

Several datasets (called recodes) are made available from each survey,
including a household recode, men’s recode and the women’s recode
(sometimes called the individual recode). This individual recode dataset
contains individual-level data on women aged 15 to 49 in the surveyed.

### Load the required Packages.

```{r}
# Load packages
 if(!require("pacman")) install.packages("pacman")
 ## Loading required package: pacman
 3
pacman::p_load(
 here,# forcermdstousetheprojectfolderasworkingdirectory
 haven,#forreadinginstatafiles
 tidyverse,#foreverything!
 here
 )
```

First, I download the dataset from the Demographic and Health
Survey(DHS)website.

### Question2.1

-   Write code below to load your data.

The dataset from DHS comes in theform of a .DTA file (the main data file
in STATA format).To import the DHS data set from the. DTA format into R
I use the read_dta() function from the {haven} package.

DHS files can be quite large–the Nigeria 2018 data has over 40,000 rows
and 5,000columns. To avoid loading all the columns that may end up
taking too long to load, I use the col_select arguement of the
read_dta() to specify which columns to import for the analysis.

The columns that we want for the analysis include(age, location,
education, etc). To check which columns to pick ,there is a data
dictionary in the data folder in the project that explains what each
variables mean.

```{r}
#Readingin the.dtafile
ir_raw <- 
  haven::read_dta(here("data/NGIR7BFL.DTA"),
 col_select= c(
 #basicinformation
 v006,v007,v009,v010,v012,v013,
 v022,v024,v025,v106,v130,v190,
 #fertility,reproductivetiming,literacylevelandmedia
 v201,v212,v213,v511,v531,
 v213,v158,v302,v155))
```

Since the DHS variable names are not very descriptive (like v000 and
v001),we will use rename() to give them short, descriptive titles.

```{r}
#Renaming thevariables
 ir_renamed <- 
  ir_raw%>%
  rename(
    Mth_interview= v006,
    Yr_interview= v007,
    Mth_birth= v009,
    Yr_birth= v010,
    Age= v012,
    Age_group= v013,
    State= v022,
    Region= v024,
    Urban_rural= v025,
    Highest_educ= v106,
    Religion= v130,
    Wealth_index= v190,
    Num_kids= v201,
    Age_first_birth= v212,
    Curr_pregnant = v213,
    Age_first_cohabitation = v511,
    Age_first_sex = v531,
    Currently_pregnant = v213,
    Listen_radio = v158,
    Ever_used_contraceptive = v302,
    Literacy = v155)
```

### Question 2.2

-   Use the colnames() to display the column names of your data set.

```{r}
colnames(ir_renamed)
```

-   Explain what the columns represent.

This data set contains 20 columns. The “Mth_interview” column represents
the monththe interview was carried out, “Yr_interview”representst the
year the interviewwas carried out, “Mth_birth” represent the month to
which the respondent was born, “Yr_birth” represent the year to which
the respondent was born, “Age” represent the age of the respondent,
“Age_group” is the age group of the respondent, “State” represent the
state where the respodent lives in “Region” represent the region where
the respondent is from, “Urban_rural” represent the type of setting to
which the respondent lives, “Highest_educ” represent the highest level
of education that the respondent has attained, “Religion” represent the
religion to which the respondent is af filiated to, “Literacy”represents
wether the respondent can read/write on not, “Listen_radio” represent
the frequency to which the respondents listens to radio, “Wealth_index”
represents the level of wealth of the respondents, “Num_kids” represents
the total number of children in a particular household,
“Age_first_birth” represents the age to which the respondent had first
child, “Currently_pregnant” represents wether the respondent is
expectant or not, “Ever_used_contraceptive” is a knowledge question if
the respondent has ever used any form of contraceptive,
“Age_first_cohabitation” represent the age at which the respondent was
married then finally age at first sex represents the age a which the
respondent had sex for the first time.

While importing the data the read_dta() function reads in factor data as
a special data type called labelled data

**Review data types.**

```{r}
glimpse(ir_renamed)
```

Several variables are of class “dbl+lbl”. This means that the columns
display only numbers, and each number corresponds to a survey answer.
This is cryptic and can be problematic for later analysis, so we convert
these to regular R factors using the function haven::as_factor(). This
will replace the numbers with the corresponding text label.

```{r}
# convert the data from labelled data to regular factors.
ir_clean <- 
  ir_renamed %>% 
  haven::as_factor()
```

Now we can see that the variables that were previously “dbl+lbl” are now
“fct” (factor) and the numbers are replaced with descriptive text.

```{r}
glimpse(ir_clean)
```

### Question 2.3

-   Work with the janitor package to clean columns names. The code below
    guide you through the process.

```{r}
library(janitor)
```

```{r}
nigeria_clean <- clean_names(ir_clean)
```

-   What transformations did the clean_names() function make to your
    data frame columns.

```{r}
colnames(nigeria_clean)
```

The clean_names() function from the janitor package helped in cleaning
the column names by changing the letters from capital to small letters
and adding an under-score fro easy analysis.

-   In your opinion, were these changes beneficial to your analysis.
    Explain.

The changes made by the clean_names() functions were beneficial to the
analysis project by changing the column names to names that are all in
lower-case, adding underscore in the column names and removing spaces in
the column names for easy manipulation.

### Question 2.4

-   2.4 A) Begin by diving into the distribution of specific variable(s)
    across a time or a date element. Use the glimpse() function and
    point out the number of rows, some data types and columns in your
    data.

The glimpse() function in R is a convenient way to get a quick overview
of the nigeria_clean data frame, showing the number of rows, data types,
and a preview of columns. To dive into the distribution of a specific
variable(s) across a time or date element,we use glimpse() function with
the nigeria_clean dataset and focus on the relevant columns which are
the num_kids and the yr_birth columns.

```{r}
#  Load the dplyr library for data manipulation.
library(dplyr)

# Use the glimpse() function to explore the data.
dplyr::glimpse(nigeria_clean)
```

```{r}
# To focus on the distribution of the variable across mth_interview and yr_interview:
glimpse(select(nigeria_clean, num_kids, yr_birth))
```

The nigeria_clean data frame has 20 columns with 41,821 rows. The data
frame also consists of mainly two data types which are double and
factors.

```{r}
summary(nigeria_clean$num_kids)
```

The summary() function shows the summary for the “num_kids” variable
giving out the minimum value of number of kids variable, the 1st
quantile, median, mean, third quantile and the maximum values.

-   2.4 B) Plot and analyze the distribution of a specific variable(s)
    across a time or a date element. Use the layers/geoms from ggplot2
    to visualize your findings. You may create 2 graphs or 1 graph and a
    summarized table where you calculate relevant statistics. You must
    include a minimun a title, the data sourse and clean up your legend,
    X and Y axis in a professional and visually pleasing manner.

The variables that I will plot for this visualization is the yr_birth
and fill against two other variables i.e “currently_pregnant” for the
pregnancy status and “literacy” ability of the respondent to read part
or full sentence or not.

```{r}
ggplot(data = nigeria_clean,
 mapping = aes(x = yr_birth,
 fill = currently_pregnant)) +
 geom_histogram(color = "black",
 position = "stack") +
 labs(title = "There are respondents below 18yrs and above 35yrs who are pregnant",
 x = "year of birth",
 y = "count",
 caption = "https://dhsprogram.com")
```

```{r}
ggplot(data = nigeria_clean,
 mapping = aes(x = yr_birth,
 fill = literacy)) +
 geom_histogram(color = "black",
 position = "stack") +
 labs(title = "Literacy level increase with decrease in age",
 x = "year of birth",
 y = "count",
 caption = "https://dhsprogram.com")
```

-   2.4 C) Identify the periods or locations in which the variables of
    interest showed significant increase, decrease or remained steady.

The variable of interest is the age of first birth and how it has been
affected over the years due to various factors including government
policy. We create a summary statistics on how the mean age of first
birth has fluctuated over a period of a ten year interval.

```{r}
mean_age1 <- nigeria_clean %>%
 filter(yr_birth %in% c(1970, 1971, 1972, 1973, 1974, 1975, 1976,
 1977, 1978, 1979)) %>%
 summarise(mean_age_first_birth = round(mean(age_first_birth, na.rm = TRUE)))
```

The mean age at first birth in from the year 1970 to 1979 was 21 years.

```{r}
mean_age2 <- nigeria_clean %>%
 filter(yr_birth %in% c(1980, 1981, 1982, 1983, 1984, 1985,
 1986, 1987, 1988,1989)) %>%
 summarise(mean_age_first_birth = round(mean(age_first_birth, na.rm = TRUE)))
```

In the period between 1980 to 1989, the mean age for first birth was 20
years which was a marginal drop compared to the mean age for first birth
between the year 1970 to 1979 which was 21 years.

```{r}
mean_age3 <- nigeria_clean %>%
 filter(yr_birth %in% c(1990, 1991, 1992, 1993, 1994,
 1995, 1996, 1997, 1998, 1999)) %>%
 summarise(mean_age_first_birth = round(mean(age_first_birth, na.rm = TRUE)))
```

In the period between 1990 to 1999, the mean age for first birth dropped
drastically from 20 years for the years between 1980 to 1989 to 18 years
for the years between 1990 to 1999.

```{r}
mean_age4 <-nigeria_clean%>%
 filter(yr_birth %in% c(2000, 2001,2002))%>%
 summarise(mean_age_first_birth= round(mean(age_first_birth, na.rm= TRUE)))
```

The mean age for first birth from the year 2000 to 2002 drops from 18
years for the years 1990 to 1999 to 16years.

```{r}
 mean_age5 <-nigeria_clean%>%
 filter(yr_birth ==2003)%>%
 summarise(mean_age_first_birth= round(mean(age_first_birth, na.rm= TRUE)))
```

For respondents born in the year 2003, the mean age at first birth is
14, which is a drastic drop from the previous years of 16.

```{r}
#Combine the mean ages into a data frame
 mean_ages <- data.frame(
 Year= c("1970-1979", "1980-1989","1990-1999","2000-2002","2003"),
 Mean_Age_First_Birth= c(mean_age1$mean_age_first_birth,
                         mean_age2$mean_age_first_birth,
                         mean_age3$mean_age_first_birth,
                         mean_age4$mean_age_first_birth,
                         mean_age5$mean_age_first_birth)
 )
 
 #Create a line graph using ggplot2
 library(ggplot2)
 ggplot(mean_ages, aes(x= Year,y= Mean_Age_First_Birth,group= 1))+
 geom_point(size= 2,
 colour="blue") +
 geom_line(size= 0.7,
 color= "brown") +
 labs(title= "MeanAgeatFirstBirthOvertheYears",
 x= "Year",
 y= "MeanAgeatFirstBirth",
 caption= "https://dhsprogram.com")
```

-   2.4 D) Discuss the potential reasons for these trends based on your
    understanding or external research, form a hypothesis you might want
    to explore in the future.

Generally over the years, there has been a drop in the mean age at which
the respondents in Nigeria have the first birth from a mean of 20.58 in
the 70s to 14 in the year 2003. This can be attributed to a number of
reasons such Economic conditions and access to opportunities. Economic
instability, limited job prospects, and poverty can incentivize early
childbearing, as young couples can perceive children as additional
sources of support. Poor access to education is also a factor as girls
who do not attend school or drop out early are more likely to have their
first child at a young age than those who complete their education.
Economic factors such as poverty can cause girls from poor households or
communities to have limited access to reproductive health services,
information, and contraception, which can increase their risk of
unintended pregnancy and early childbearing. Some Nigerian communities
may have traditions or beliefs that encourage or pressure girls to have
children early, either for religious, economic, or social reasons. These
norms may also discourage girls from using contraception or seeking
abortion if they become pregnant before they are ready.

-   Hypothesis.

The remarkable reduction in the mean age at first childbirth in Nigeria,
signifying early childbearing among young individuals, may be attributed
to a complex interplay of socioeconomic, cultural, educational, health
care, policy, and societal factors. Future research should undertake a
multifaceted investigation to quantify the specific contributions of
these factors in driving the trend, explore regional variations within
Nigeria, and identify potential intervention points to address early
childbearing among the youth population.

-   2.4 F) Discuss these trends in relation to your original research
    question. Does this preliminary analysis hint towards any potential
    answers or raise more questions.

One of our research questions was that the Federal Government of Nigeria
has placed policies to ensure that the minimum age for pregnancy be 18
years. From our preliminary analysis, we found out that the mean age of
the respondents first pregnancy over the years has fallen from 20 years
in the 70s to 14 years in the year 2003. This is way below the required
minimum standards set by the government to realize a reduced population
growth rate in Nigeria.

## Graded Task 3: Modelling

After diving into the data exploration, its time to operationalize our
insights. This section requires you to decide on a dependent variable
(Y) and an independent variable (X). Ensure that these variables are
pertinent to your research questions.

### Question 3.1

-   what did you select as your independent and dependent variables.

My Dependent Variable (Y) for this analysis is the “num_kids” variable
and the independent variable (X) for this analysis are: 1.
“age_first_cohabitation”. 2. “age_first_birth”. 3. “listen_radio”. 4.
“literacy”. 5. “religion”. 6. “highest_educ”.

-   Explain the rationale behind your choice, how does these variables
    relate to or help answer your research questions?

The rationale behind selecting these independent variables is that they
represent various factors that could potentially impact birth rates and,
consequently, population growth. By analyzing these variables in
relation to the number of children in a household, we can assess whether
government policies aimed at reducing population growth have had an
effect, and whether sociodemographic factors such as age at first
cohabitation, age at first birth, exposure to public education through
radio, literacy levels, religion, and education level play a role in
influencing birth rates in Nigeria.

### Question 3.2

How would you write out your regression formula for your Y and X
variable.

Recall the regression formula 𝑌 −𝐵0+𝐵1+𝐶 where • 𝑌 is your
expected/predicted outcome, if 𝑌 was wage then it would be expected_wage
• 𝐵0 is the intercept. It represents the expected value of 𝑌 when X is 0
• 𝐵1 is the slope. It indicates the change in 𝑌 for a one unit change in
X variable.

num_kids- 6.5304 + (-0.1342) + C ( for the “age_first_cohabitation”
variable) Question 3.3) Now that you have written out your formula,
program your regression in R and show your output using the summary
function.

```{r}
 lm1 <- lm(formula = num_kids ~ age_first_cohabitation, data = nigeria_clean)
 lm2 <- lm(formula = num_kids ~ listen_radio, data = nigeria_clean)
 lm3 <- lm(formula = num_kids ~ literacy, data = nigeria_clean)
 lm4 <- lm(formula = num_kids ~ religion, data = nigeria_clean)
 lm5 <- lm(formula = num_kids ~ highest_educ, data = nigeria_clean)
```

We can now run the summary of each model.

```{r}
summary((lm1))
```

The multiple R-squared value (0.06053) indicates the proportion of
variability in the birth rate that can be explained by the age at first
cohabitation. In this case, approximately 6.05% of the variability in
birth rate is explained by this model. The F-statistic (2007) and its
associated p-value (\< 2.2e-16) test whether the overall model is
statistically significant. In this case, the p-value is less than 0.05
suggests that the model as a whole is statistically significant. In
summary, the “lm1” model suggests that there is a statistically
significant negative relationship between the age at first cohabitation
and the birth rate in Nigeria. As the age at first cohabitation
decrease, the number of kids / birth rate tends to increase, while
controlling for other factors in the model.

```{r}
summary(lm2)
```

The multiple R-squared value (0.004769) indicates the proportion of
variability in the birth rate that can be explained by the radio
listening habits (independent variable). In this case, radio listening
habits explain only a very small portion (less than 1%) of the
variability on birth rate. The F-statistic (100.2) and its associated
p-value of \< 2.2e-16 which is less than 0.05 in this case, shows that
the model is statistically significant. In summary, the “lm2” model
suggests that there is a statistically significant relationship between
radio listening habits and the birth rate in Nigeria. For example,
respondents who listen to radio at least once a week is associated with
a lower birth rate by approximately 0.3361 units compared to respondents
who listen to radio less than once a week.

```{r}
summary((lm3))
```

The multiple R-squared value (0.1418) indicates the proportion of
variability in the birth rate that can be explained by the literacy
levels. In this case, approximately 14.18% of the variability in birth
rates is explained by this model. The F-statistic (1728) and its
associated p-value (\< 2.2e-16) which is also less than 0.05 shows that
the model as a whole is highly statistically significant. In summary,
the “lm3” model suggests that there is a statistically significant
relationship between literacy levels and the birth rate in Nigeria. As
individuals’ literacy levels decrease, the birth rate tends to increase,
holding other factors constant.

```{r}
summary(lm4)
```

The multiple R-squared value (0.03934) indicates the proportion of
variability in the birth rate that can be explained by individuals’
religious affiliation (religion variable). In this case, approximately
3.93% of the variability in birth rates is explained by this model. The
F-statistic (428.1) and its associated p-value of (\< 2.2e-16) which is
less than 0.05 shows that the model as a whole is highly statistically
significant. In summary, the “lm4” model suggests that there is a
statistically significant relationship between individuals’ religious
affiliation and the birth rate in Nigeria. The type of religion
individuals identify with is associated with differences in birth rates,
with “islam” and “traditionalist” categories being associated with
higher birth rates compared to the reference category.

### Question 3.3

Now time to program a multi-variate regression using an additional X. I
dentify a second X that you want to include and,

A)  ) Rewrite your regression formula to include this additional
    variable. • birth_rate- 6.5304 + (-0.1342) + C- Co-efficient(X2) \*
    X2 • birth_rate- 6.5304 + (-0.1342) + C- 4.4242 \* highest_educ (for
    “age_first cohabitation” and “high est_educ” variables).

WHERE; • birth_rate: The dependent variable • 6.5304: The intercept
(constant term). •-0.1342: The coefficient for the first independent
variable. • C: Any constant. • 4.4242: The coefficient for the second
independent variable, which is represented by X2.

B)  ) Re-run your regression including your new X and provide an
    interpretation of your outcomes, include a statistical
    interpretation on your model model fit.

```{r}
lm6 <- lm(formula = num_kids ~ age_first_cohabitation + highest_educ, data = nigeria_clean)
```

```{r}
summary(lm6)
```

Multiple R-squared (0.1046) represents the proportion of variance in the
dependent variable explained by the model. Adjusted R-squared (0.1045)
adjusts for the number of predictors in the model. This statistic tests
the overall significance of the model. In this case, the F-statistic is
909.7 with a very low p-value (\<2.2e-16) less than 0.05, indicates that
the model as a whole is statistically significant. Overall, this
regression model suggests that age_first_cohabitation and the level of
education (highest_educ) are associated with changes in birth_rate, with
individuals whose age at first cohabitation higher and have secondary or
higher education having lower birth rates. Extra Credit Question 3.6)
Now we ask that you take your regression outputs and improve their
readabilty by using arguments in the stargaze package, include both
models. ( Hint: I showed this package in week 5, see the rmd file ).
Please note the” result as is ” chunk option and you will see the pretty
table in the final knitted PDF and not in the chunk output. Your model
name should be according to what they are for example, BI-variate or
Multi-variate, dive into the stargazer documentation to make it as nice
as possible.

```{r}
if(!require(pacman)) install.packages("pacman")
pacman::p_load(stargazer)
  
```

```{r}
# Create a list of models with names
 my_models <- list(
   "Bivariate Model4" = lm4,
   "Multivariate Model6" = lm6
 )
 # Generate the table with named models
 stargazer::stargazer(my_models,
 title = "2 Binary Models",
 type = "text",
 keep.stat = c("n", "rsq", "f"),
 column.labels = c("Bivariate Model4", "Multivariate Model6"),
 header = FALSE) # Prevents the addition of LaTeX headers
```

**HWExtra Credit: Fixed Effects** Gain some extra points by delving into
the world of regression modelling. Note: Accurately answering each
question will earn you 1 point returned on a previous homework of your
choice. Please specify the homework assignment where you’d like the
point to be added.

**EC Question 1)** As mentioned, pick a time or spacial element you want
to control for in your regression analysis. This is often important in
empirical research to capture unobserved heterogenity. • Which specific
element ( e.g., year, region, etc) did you choose for and why?

Controlling for the “region” is an appropriate choice in regression
analysis, as it can help capture spatial variations and unobserved
heterogeneity in the data. By including “region” as a fixed effect, we
can account for differences in outcomes that might be specific to
different regions. This can be important, especially in the context of
government policies and their impact on population growth, as regions
may have distinct policy implementations or social and economic
characteristics.

The “region” variable could confound the relationship between the
dependent variable (e.g Number of children) and the independent variable
(e.g., age at first cohabitation) because different regions may have
varying cultural, economic, and social factors that influence birth
rates differently. For example, one region might have a higher literacy
rate and better access to family planning services, resulting in lower
birth rates, while another region might have different cultural norms or
economic conditions that lead to higher birth rates. By including
“region” as a fixed effect, we control for these unobserved regional
variations, making the analysis more robust and allowing us to focus on
the specific impact of the independent variable on the dependent
variable across regions. • How might this element confound the
relationship between your department and independent variable?

Regions can have unique characteristics, policies, and cultural factors
that influence both the dependent and independent variables differently.
For example, a region with strong family planning initiatives may have a
lower birth rate, which can confound the relationship between the age at
first cohabitation and number of kids. The effect of age at first
cohabitation on birth rate may differ across regions. Again, Including
“region” as a fixed effect may introduce multicollinearity into the
regression model. Multicollinearity occurs when two or more independent
variables are highly correlated, making it challenging to discern the
individual effects of each variable. If the region has a strong
influence on both the dependent and independent variables, it can lead
to multicollinearity and make it difficult to isolate the specific
impact of the independent variable.

**EQ Question 2)**

Fixed effects can be used to control for time invariant unobserved
heterogenity. Implement a fixed- effects regression in R, controlling
for the element you selected in EC Question 1. Show your code and output
using summary() function.

```{r}
 # Load the plm package
 if (!require("pacman")) install.packages("pacman")
pacman::p_load(plm)
```

```{r}
# Create a panel data frame with fixed effects
panel_data <- pdata.frame(nigeria_clean, index = c("region"))
 # Run a fixed-effects regression
 lm7 <- plm(num_kids ~ age_first_cohabitation + highest_educ, data = panel_data, model = "within")
 # Summarize the model
 summary(lm7)
```

**EC Question 3**

Improve your readability by including your model into the previous
pretty stargazer table you made from **Question3.5** and show us your
updated table, be sure to name the models appropriately.

```{r}
#Create a listofmodelswithnames
 my_models <-list(
 "BivariateModel4" = lm4,
 "MultivariateModel6" = lm6,
 "fixed_model"= lm7
 )
 #Generate thetablewithnamedmodels
 stargazer::stargazer(my_models,
 title= "2BinaryModels",
 type= "text",
 keep.stat= c("n","rsq", "f"),
 column.labels = c("BivariateModel4", "MultivariateModel6", "fixedeffects"),
 header= FALSE) #PreventstheadditionofLaTeXheaders
```

**EC Question 4)** Compare the results of this fixed effects regression
with your previous multivariate regression.

-   Have the co-efficients on your main variable of interest changed
    significantly?

The coefficient for “age_first_cohabitation” has changed slightly
from-0.079615 in the multivariate regression to-0.0820391 in the
fixed-effects regression. While this change is relatively small, it
indicates that accounting for the region as a fixed effect may have a
slight influence on the relationship between age at first cohabitation
and the number of children. This could be due to regional variations in
unobserved factors that affect the dependent variable.

-   What can you infer about the potential confounding influence of the
    time or spacial element you controlled for?

The inclusion of “region” as a fixed effect in the fixed-effects
regression is intended to control for spatial variations. This means
that we are accounting for differences in outcomes that may be specific
to different regions. The slightly different coefficient for
“age_first_cohabitation” in the fixed-effects regression suggests that
regional differences may indeed influence the relationship between age
at first cohabitation and the number of children. While controlling for
region helps reduce spatial confounding, there might still be unobserved
regional specific factors affecting the relationship. It is essential to
acknowledge that even after using fixed effects, unobserved confounding
variables related to both time and space may persist, but the model has
become more robust by accounting for region-specific variations.

**References**

-   Okpala, A. O. (1990). Nigerian population growth and its
    implications for economic development. Scandinavian Journal of
    Development Alternatives, 9(4), 63-77.

-   Mazzocco, K. (1988). Nigeria’s new population policy. International
    health news, 9(3), 1-12.

-   <https://dailypost.ng/2021/05/14/fg-must-take-urgent-steps-to-address-rapid-population-growth>
    health-minister-osagie-ehanire/

-   <https://dhsprogram.com/data/new-user-registration.cfm>
